---
title: "simulation"
date: "12/13/2019"
output: pdf_document
---

```{r setup, include=FALSE, echo=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(tidy = TRUE)
library(tidyverse)
library(lfe)
library(lme4)
library(lmerTest)
library(gridExtra)
iter <- 100
```

Section I: Base Scenario for Comparison (Meets Assumptions)

```{r}
source("dgp_script.R") #all data generation processes are here
set.seed(2020)
classroom_dat <- class_dat_function(40, 25)
```

```{r}
par(mfrow = c(2,3))
plot(classroom_dat$yearstea, classroom_dat$avgtest, xlab = "Years teaching", ylab="Average Test Scores", main = "")
hist(classroom_dat$fam.income, main = "Family Income")
hist(classroom_dat$pretest, main = "Pretest Scores")
plot(classroom_dat$parent.edu, classroom_dat$fam.income, xlab= "Parent Education", ylab = "Family Income")
plot(classroom_dat$pretest, classroom_dat$Y_stud, xlab = "9th grades scores", ylab = "10th grade scores", main = "Randomized at student-level")
plot(classroom_dat$pretest, classroom_dat$Y_class, xlab = "9th grades scores", ylab = "10th grade scores", main = "Randomized at classroom-level") 
```

```{r}
ggplot(data=classroom_dat, aes(x=classid, y = Y_stud)) + geom_boxplot() + 
  theme_classic() +labs(x = "Class ID", y = "10th grade test scores", title = "Randomization at student level")
```

```{r}
classroom_dat %>% mutate(treatment = ifelse(Z_class == 1, "Treatment", "Control")) %>% 
  ggplot(aes(x = classid, y = Y_class, fill = treatment)) + geom_boxplot() + 
  theme_classic() +labs(x = "Class ID", y = "10th grade test scores", title = "Randomization at classroom level")
```


```{r}
SATE <- mean(classroom_dat$Y1) - mean(classroom_dat$Y0)
model_stu <- classroom_dat %>% select(Y_stud, Z_stud, yearstea, teach.edu, avgtest,
                                      minority, parent.edu, fam.income, freelunch, dist.school.hour, 
                                      gender, pretest, classid)

#Base Linear Regression
lr.base <- lm(Y_stud ~., data = model_stu[,-13])
summary(lr.base)

#Base - Linear Regression Model Bias
(lr.base.bias <- lr.base$coefficients[2] - SATE)

#Base w/ Fixed Effects
fixed.base <- lfe::felm(model_stu$Y_stud ~ model_stu$Z_stud + model_stu$minority + model_stu$parent.edu +
                          model_stu$fam.income + model_stu$freelunch + model_stu$dist.school.hour + model_stu$gender +
                          model_stu$pretest | model_stu$classid) 
summary(fixed.base)

fixed.base.factor <- lm(model_stu$Y_stud ~ model_stu$Z_stud + model_stu$minority + model_stu$parent.edu +
                          model_stu$fam.income + model_stu$freelunch + model_stu$dist.school.hour + model_stu$gender +
                          model_stu$pretest + factor(model_stu$classid))
summary(fixed.base.factor)$coefficients[2]

#Base - Fixed Effects Model Bias
(fe.base.bias <- fixed.base$coefficients[1] - SATE)

#Base w/ Random Effects
random.base <- lmerTest::lmer(Y_stud ~ Z_stud + minority + parent.edu + fam.income + freelunch + dist.school.hour +
                                gender + pretest + yearstea + avgtest + teach.edu + (1 | classid), data = model_stu)
summary(random.base)

#Base - Random Effects Model Bias
(re.base.bias <- summary(random.base)$coefficients[2,1] - SATE)

```

```{r, message=FALSE}
# Randomization Distribution - Base Case

base_rd_sim <- data.frame(type = rep(NA, iter*3), coef = rep(NA, iter*3), 
                          conf_int_low = rep(NA, iter*3), conf_int_high = rep(NA, iter*3))
count <- 1

for(i in 1:iter){
  
  #Randomization Distribution - randomize treatment
  base_data <- classroom_dat
  N <- nrow(base_data)
  
  X_stud <- rnorm(N, 0, 1)
  prob_stud <- inv.logit((X_stud/max(abs(X_stud))) * log(19))
  Z_stud <- rbinom(N, 1, prob = prob_stud)
  
  base_data$Z_stud <- Z_stud
  base_data$Y_stud <- ifelse(Z_stud == 1, base_data$Y1, base_data$Y0)
  
  base_data_model <- base_data %>% select(Y_stud, Z_stud, yearstea, teach.edu, avgtest, minority, parent.edu, 
                                          fam.income, freelunch, dist.school.hour, gender, pretest, classid)

  #Linear Regression Sim
  j <- count
  base_rd_sim[j,1] <- "lr"
  lr_base_sim <- lm(Y_stud ~., data = base_data_model[,-13])
  base_rd_sim[j,2] <- lr_base_sim$coefficients[2]
  base_rd_sim[j,3] <- confint(lr_base_sim, 'Z_stud', level = .95)[1,1]
  base_rd_sim[j,4] <- confint(lr_base_sim, 'Z_stud', level = .95)[1,2]
  
  #Fixed Effect Sim
  j <- count+1
  base_rd_sim[j,1] <- "fixed"
  fixed_base_sim <- lfe::felm(base_data_model$Y_stud ~ base_data_model$Z_stud + base_data_model$minority +
                            base_data_model$parent.edu + base_data_model$fam.income + base_data_model$freelunch +
                            base_data_model$dist.school.hour + base_data_model$gender + base_data_model$pretest |
                            base_data_model$classid) 
  base_rd_sim[j,2] <- fixed_base_sim$coefficients[1]
  base_rd_sim[j,3] <- confint(fixed_base_sim, 'base_data_model$Z_stud', level = .95)[1,1]
  base_rd_sim[j,4] <- confint(fixed_base_sim, 'base_data_model$Z_stud', level = .95)[1,2]
  
  #Random Effect Sim
  j <- count+2
  base_rd_sim[j,1] <- "random"
  random_base_sim <- lmerTest::lmer(Y_stud ~ Z_stud + minority + parent.edu + fam.income + freelunch + dist.school.hour +
                                gender + pretest + yearstea + avgtest + teach.edu + (1 | classid), data = base_data_model) 
  base_rd_sim[j,2] <- summary(random_base_sim)$coefficients[2,1]
  base_rd_sim[j,3] <- confint(random_base_sim, 'Z_stud', level = .95)[1,1]
  base_rd_sim[j,4] <- confint(random_base_sim, 'Z_stud', level = .95)[1,2]  
  
  count <- j+1
  print(i)
}

base_rd_sim$SATE <- rep(SATE, nrow(base_rd_sim))

```

```{r}

hist(base_rd_sim$coef[base_rd_sim$type == "lr"], main = "Randomization Distribution - Base (Linear Regression)", xlab = "Treatment Effect", xlim = c(min(base_rd_sim$coef)-.5, max = max(base_rd_sim$coef)+.5))
abline(v = SATE, col = "red")

hist(base_rd_sim$coef[base_rd_sim$type == "fixed"], main = "Randomization Distribution - Base (Fixed Effects)", xlab = "Treatment Effect", xlim = c(min(base_rd_sim$coef)-.5, max = max(base_rd_sim$coef)+.5))
abline(v = SATE, col = "red")

hist(base_rd_sim$coef[base_rd_sim$type == "random"], main = "Randomization Distribution - Base (Random Effects)", xlab = "Treatment Effect", xlim = c(min(base_rd_sim$coef)-.5, max = max(base_rd_sim$coef)+.5))
abline(v = SATE, col = "red")

```

```{r, message=FALSE}
# Sampling Distribution - Base Case

base_sd_sim <- data.frame(type = rep(NA, iter*3), coef = rep(NA, iter*3), 
                          conf_int_low = rep(NA, iter*3), conf_int_high = rep(NA, iter*3), SATE = rep(NA, iter*3))
count <- 1

for(i in 1:iter){
  
  #Sampling Distribution - resample data
  base_data <- class_dat_function(40, 25)
  SATE_sim <- mean(base_data$Y1) - mean(base_data$Y0)
  base_data_model <- base_data %>% select(Y_stud, Z_stud, yearstea, teach.edu, avgtest, minority, parent.edu, 
                                          fam.income, freelunch, dist.school.hour, gender, pretest, classid)

  #Linear Regression Sim
  j <- count
  base_sd_sim[j,1] <- "lr"
  lr_base_sim <- lm(Y_stud ~., data = base_data_model[,-13])
  base_sd_sim[j,2] <- lr_base_sim$coefficients[2]
  base_sd_sim[j,3] <- confint(lr_base_sim, 'Z_stud', level = .95)[1,1]
  base_sd_sim[j,4] <- confint(lr_base_sim, 'Z_stud', level = .95)[1,2]
  base_sd_sim[j,5] <- SATE_sim

  #Fixed Effect Sim
  j <- count+1
  base_sd_sim[j,1] <- "fixed"
  fixed_base_sim <- lfe::felm(base_data_model$Y_stud ~ base_data_model$Z_stud + base_data_model$minority +
                            base_data_model$parent.edu + base_data_model$fam.income + base_data_model$freelunch +
                            base_data_model$dist.school.hour + base_data_model$gender + base_data_model$pretest |
                            base_data_model$classid) 
  base_sd_sim[j,2] <- fixed_base_sim$coefficients[1]
  base_sd_sim[j,3] <- confint(fixed_base_sim, 'base_data_model$Z_stud', level = .95)[1,1]
  base_sd_sim[j,4] <- confint(fixed_base_sim, 'base_data_model$Z_stud', level = .95)[1,2]
  base_sd_sim[j,5] <- SATE_sim

  #Random Effect Sim
  j <- count+2
  base_sd_sim[j,1] <- "random"
  random_base_sim <- lmerTest::lmer(Y_stud ~ Z_stud + minority + parent.edu + fam.income + freelunch + dist.school.hour +
                                gender + pretest + yearstea + avgtest + teach.edu + (1 | classid), data = base_data_model) 
  base_sd_sim[j,2] <- summary(random_base_sim)$coefficients[2,1]
  base_sd_sim[j,3] <- confint(random_base_sim, 'Z_stud', level = .95)[1,1]
  base_sd_sim[j,4] <- confint(random_base_sim, 'Z_stud', level = .95)[1,2]  
  base_sd_sim[j,5] <- SATE_sim

  count <- j+1
  print(i)
}

```

```{r}

hist(base_sd_sim$coef[base_sd_sim$type == "lr"], main = "Sampling Distribution - Base (Linear Regression)", xlab = "Treatment Effect", xlim = c(min(base_sd_sim$coef)-.5, max = max(base_sd_sim$coef)+.5))
abline(v = mean(base_sd_sim$SATE), col = "red")

hist(base_sd_sim$coef[base_sd_sim$type == "fixed"], main = "Sampling Distribution - Base (Fixed Effects)", xlab = "Treatment Effect", xlim = c(min(base_sd_sim$coef)-.5, max = max(base_sd_sim$coef)+.5))
abline(v = mean(base_sd_sim$SATE), col = "red")

hist(base_sd_sim$coef[base_sd_sim$type == "random"], main = "Sampling Distribution - Base (Random Effects)", xlab = "Treatment Effect", xlim = c(min(base_sd_sim$coef)-.5, max = max(base_sd_sim$coef)+.5))
abline(v = mean(base_sd_sim$SATE), col = "red")

ggplot() +
  geom_histogram(data = base_sd_sim[base_sd_sim$type == "lr",], aes(x = coef), color = "blue", fill = "white", alpha = .25) +
  geom_histogram(data = base_sd_sim[base_sd_sim$type == "fixed",], aes(x = coef), color = "green", fill = "white", alpha = .25) +
  geom_histogram(data = base_sd_sim[base_sd_sim$type == "random",], aes(x = coef), color = "red", fill = "white", alpha = .25) +
  labs(x = "Treatment Effect", y = "Frequency") +
  geom_vline(xintercept=SATE, color = "black", linetype = "dashed", size = 1.25)

```



Section II: Violate Random Effects Assumption
```{r}
re_dat <- re_dat_function(40, 25)

SATE_re <- mean(re_dat$Y1) - mean(re_dat$Y0)

model_stu_re <- re_dat %>% select(Y_stud, Z_stud, yearstea, teach.edu, avgtest, minority, parent.edu, 
                                  fam.income, freelunch, dist.school.hour, gender, pretest, classid)

#Random Effects Violation Linear Regression
lr.re <- lm(Y_stud ~., data = model_stu_re[,-13])
summary(lr.re)

#Random Effects Violation - Linear Regression Model Bias
(lr.re.bias <- lr.re$coefficients[2] - SATE_re)

#Random Effects Violation w/ Fixed Effects
fixed.re <- lfe::felm(model_stu_re$Y_stud ~ model_stu_re$Z_stud + model_stu_re$minority + model_stu_re$parent.edu +
                        model_stu_re$fam.income + model_stu_re$freelunch + model_stu_re$dist.school.hour +
                        model_stu_re$gender + model_stu_re$pretest | model_stu_re$classid) 
summary(fixed.re)$coefficients[1]

#Random Effects Violation - Fixed Effects Model Bias
(fe.re.bias <- fixed.re$coefficients[1] - SATE_re)

#Random Effects Violation w/ Random Effects
random.re <- lmerTest::lmer(Y_stud ~ Z_stud + minority + parent.edu + fam.income + freelunch + dist.school.hour +
                              gender + pretest + yearstea + avgtest + teach.edu + (1 | classid), data = model_stu_re)
summary(random.re)

#Random Effects Violation - Random Effects Model Bias
(re.re.bias <- summary(random.re)$coefficients[2,1] - SATE_re)

```

Section III: Violate Ignorability Assumption 
```{r}
ig_dat <- ig_dat_function(40, 25)

SATE_ig <- mean(ig_dat$Y1) - mean(ig_dat$Y0)

model_ig <- ig_dat %>% select(Y_stud, Z_stud, yearstea, teach.edu, avgtest, minority, parent.edu, fam.income, 
                              freelunch, gender, pretest, classid)

#Ignorability Violation - Linear Regression
lr.ig <- lm(Y_stud ~., data = model_ig[,-12])
summary(lr.ig)

#Ignorability Violation - Linear Regression Model Bias
(lr.ig.bias <- lr.ig$coefficients[2] - SATE_ig)

#Ignorability Violation w/ Fixed Effects
fixed.ig <- lfe::felm(model_ig$Y_stud ~ model_ig$Z_stud + model_ig$minority + model_ig$parent.edu + 
                        model_ig$fam.income + model_ig$freelunch + model_ig$gender + model_ig$pretest |model_ig$classid)
summary(fixed.ig)$coefficients[1]

#Ignorability Violation - Fixed Effects Model Bias
(fe.ig.bias <- fixed.base$coefficients[1] - SATE_ig)

#Ignorability Violation w/ Random Effects
random.ig <- lmerTest::lmer(Y_stud ~ Z_stud + minority + parent.edu + fam.income + freelunch + gender + pretest + 
                              yearstea + avgtest + teach.edu + (1 | classid), data = model_ig)
summary(random.ig)

#Base - Random Effects Model Bias
(re.ig.bias <- summary(random.ig)$coefficients[2,1] - SATE_ig)
```

Section IV: Treatment Assignment at the Group Level

*Make sure to interpret bias & treatment effect coefficients at the GROUP LEVEL - estimand has changed; treatment effect of the treated group*

```{r}
#Treatment Effect of the Treated @ Group Level
model_class <- classroom_dat %>% select(Y_class, Z_class, yearstea, teach.edu, avgtest, minority, parent.edu,
                                        fam.income, freelunch, dist.school.hour, gender, pretest, classid)

#Group Level Treatment Effect Linear Regression # ignore group structure
lr.gl <- lm(Y_class ~., data = model_class[,-13])
summary(lr.gl)

#Group Level Treatment Effect - Linear Regression Model Bias
(lr.gl.bias <- lr.gl$coefficients[2] - SATE)

#Group Level Treatment Effect w/ Fixed Effects
fixed.gl <- lfe::felm(model_class$Y_class ~ model_class$Z_class + model_class$minority + model_class$parent.edu +
                        model_class$fam.income + model_class$freelunch + model_class$dist.school.hour +
                        model_class$gender + model_class$pretest | model_class$classid)
summary(fixed.gl)

### 
fixed.gl.factor <- lm(model_class$Y_class ~ model_class$Z_class + model_class$minority + model_class$parent.edu +
                        model_class$fam.income + model_class$freelunch + model_class$dist.school.hour +
                        model_class$gender + model_class$pretest + factor(model_class$classid) +
                        model_class$yearstea + model_class$avgtest + factor(model_class$teach.edu))
summary(fixed.gl.factor)$coefficients[2]


#Group Level Treatment Effect - Fixed Effects Model Bias
## Z_class does not return coeffient - I think this makes sense since we're running the regression at the class level?
# (fe.gl.bias <- fixed.gl$coefficients[1] - SATE)

#Group Level Treatment Effect w/ ****Random Effects***
random.gl <- lmerTest::lmer(Y_class ~ Z_class + minority + parent.edu + fam.income + freelunch + dist.school.hour +
                              gender + pretest + yearstea + avgtest + teach.edu + (1 | classid), data = model_class)
summary(random.gl)

#Group Level Treatment Effect - Random Effects Model Bias
(re.gl.bias <- summary(random.gl)$coefficients[2,1] - SATE)

```

Comparison Tables

```{r, fig.height = 2, fig.width = 7}

bias_compare <- data.frame("Bias Comparison" = c("Linear Regression", "Fixed Effects", "Random Effects"), 
                           "Base Case" = round(c(lr.base.bias, fe.base.bias, re.base.bias),2), 
                           "RE Violation" = round(c(lr.re.bias, fe.re.bias, re.re.bias),2), 
                           "Ig. Violation" = round(c(lr.ig.bias, fe.ig.bias, re.ig.bias),2), 
                           "Group Treatment" = c(round(lr.gl.bias,2), "N/A", round(re.gl.bias,2)))

grid.arrange(tableGrob(bias_compare, theme=ttheme_default()), nrow = 1)
```


